
<!-- ```{r, message = FALSE} -->
<!-- library(tidyverse) -->
<!-- library(here) -->
<!-- source(here("./scripts/funcs.R")) -->
<!-- source(here("./scripts/plot_funcs.R")) -->
<!-- dics <- create_conversion_dics() -->
<!-- ``` -->


# Community survey

Mainly contains information about access to services, field-labour price, seed and input prices.

## Intervening datasets

```{r, echo = FALSE}
datasets <- list_datasets("03-community.Rmd")
datasets$Description <- c("Community main file",
                          "Community metadata")
knitr::kable(datasets)
```


## Prepare dataset

```{r}
cmty <- read.table(here("./data/TAMASA2017/TZAPS17_cmty.tab"),
                   header = TRUE,
                   sep = "\t",
                   stringsAsFactors = FALSE)
```

```{r}
cmty$village[cmty$village == "OTHER"] <- cmty$village_alt[cmty$village == "OTHER"]
cmty$hamlet[cmty$hamlet == "OTHER"] <- cmty$hamlet_alt[cmty$hamlet == "OTHER"]
```


Set zone and region as factors with legible levels

```{r}
md_geo <- readxl::read_xlsx(here("./data/TAMASA2017/metadata/TZAPS17_ODK_community.xlsx"),
                            sheet = "external_choices")
md_region  <- md_geo %>% filter(list_name == "regions")
md_district  <- md_geo %>% filter(list_name == "districts")
```


```{r}
cmty$zone <- as.factor(cmty$zone)
cmty$region <- factor(cmty$region,
                      levels = md_region$name,
                      labels = md_region$label)
cmty$district <- factor(cmty$district,
                        levels = md_district$name,
                        labels = md_district$label)
```

Decriptive variables:

* `cmty_index`: community index, field with only unique values: ``r length(cmty$cmty_index) == length(unique(cmty$cmty_index))``.
* `enum_date`: date of the survey.
* `zone > region > district > site > ward > village > hamlet`: spatial hierachy.
* `comm_gps`: community GPS coordinates.

Quantitative variables :

* `vil_nhh`: number of households in the village.
* `vil_npersons`: number of persons in the village.
* `vil_acres`: approximate size of the village in acres.

* `acc_inp_km`: distance to input dealer in km.
* `acc_inp_tsh`: cost of transport in Tsh.


```{r}
cmty_info <- cmty %>% 
  select(cmty_index, enum_date, zone, region, district, site, ward, village,
         hamlet, comm_gps, vil_nhh, vil_npersons, vil_acres, acc_inp_km, acc_inp_tsh)
```

Data is available for ``r nrow(cmty_info)`` communities threw out the country.

The quantitative variables should be strictly positive. Scan for negative values:

```{r}
(neg_val <- cmty_info[map_lgl(cmty_info, is.numeric)] %>% any_negative())
```

and replace existing ones by NA. 

```{r}
cmty_info[neg_val] <- modify(cmty_info[neg_val], ~ ifelse(.x < 0, NA, .x))
```


```{r}
plot_var(cmty_info, vil_nhh, plot.unit = "nb of household")
```

```{r, warning = FALSE}
plot_var_matrix(df = cmty_info,
                vars = c("vil_nhh", "vil_npersons", "vil_acres"),
                plot.unit = c("nb of household", "nb of persons", "acres"),
                plot.dim = c(1, 3))
```

```{r}
cmty_info <- modify_at(.x = cmty_info,
                       .at = c("vil_nhh", "vil_npersons", "vil_acres"),
                       .f = outlier_to_NA,
                       quantile = 0.95)
```


```{r}
del_NA <- cmty_info %>%
  select(starts_with("vil_")) %>%
  map(~ which(is.na(.x))) %>%
  unlist() %>%
  unique()
cmty_info <- cmty_info[-del_NA, ]
```


```{r}
dim(cmty_info)
```


```{r, warning = FALSE}
plot_var_matrix(df = cmty_info,
                vars = c("vil_nhh", "vil_npersons", "vil_acres"),
                plot.unit = c("nb of household", "nb of persons", "acres"),
                plot.dim = c(1, 3))
```

```{r, warning = FALSE}
plot_var_matrix(
  df = cmty_info,
  vars = c("acc_inp_km", "acc_inp_tsh"),
  plot.unit = c("km", "tsh"),
  plot.dim = c(1, 2)
)

```


A number of communities are said to be located at a 0 km from an input dealer or to have null input transportation cost. It can be that these communities are indeed located nearby a supplier or most likely that they simply do not use additional inputs.


## Seed buying prices

The community datasets contains a varierty of variables related to maize prize.
Those 9 variables, all prefixed by `maiprc`, provide the price in Tsh of a 50 kg bag of maize seeds at two possible locations (`depot` or `town`) in four different time periods (`q1`, `q2`, `q3` and `q4`).

**Locations:**

`depot` refers to a large commercial depot compare to `town` which is supposed to be the regular market place gor example in the district town.

**Time periods:**

| Buying place | Apr-Jun 2016 | Jul-Sep 2016 | Oct-Dec 2016 | Jan-Mar 2017 |
|--------------|--------------|--------------|--------------|--------------|
| depot        | q1           | q2           | q3           | q4           |
| town         | q1           | q2           | q3           | q4           |


```{r}
cmty_seed <- cmty %>% 
  select(cmty_index, starts_with("maiprc_"))
```

```{r}
maiprc <- grep("^maiprc", colnames(cmty_seed), value = TRUE)
```

Is there any price variable wiht a negative price ?

```{r}
cmty_seed[, maiprc] %>% any_negative()
```

```{r}
cmty_seed[, maiprc] <- modify(cmty_seed[, maiprc], ~ ifelse(.x < 0, NA, .x))
```

Convert price from Tsh/50 kg bag to Tsh/kg

```{r}
cmty_seed[, maiprc] <- cmty_seed[, maiprc] / 50
```

```{r, warning = FALSE}
plot.unit = "Tsh/kg"
title = "Maize price"
cmty_seed %>%
  pivot_longer(cols = maiprc,
               names_to = "category",
               values_to = "value") %>%
  separate(col = category, into = c("maiprc", "place", "trimester"), sep = "_") %>%
  plot_var(var = value, plot.unit = plot.unit)+
  facet_grid(place ~ trimester)+
  xlab("")+
  ggtitle(title)
```

Replace extreme values by NA

```{r}
cmty_seed <- modify_at(.x = cmty_seed,
                       .at = maiprc,
                       .f = outlier_to_NA,
                       quantile = 0.99)
```

and delete any community harbouring NA for any variable

```{r}
del_NA <- cmty_seed %>%
  map(~ which(is.na(.x))) %>% 
  unlist() %>%
  unique()
cmty_seed <- cmty_seed[-del_NA, ]
```


```{r, warning = FALSE}
plot.unit = "Tsh/kg"
title = "Maize price"
cmty_seed %>%
  pivot_longer(cols = maiprc,
               names_to = "category",
               values_to = "value") %>%
  separate(col = category, into = c("maiprc", "place", "trimester"), sep = "_") %>%
  plot_var(var = value, plot.unit = plot.unit)+
  facet_grid(place ~ trimester)+
  xlab("")+
  ggtitle(title)
```


## Fertilizer prices


### Selected variables

#### Standard bag 

Cost in Tsh of a 50 kg bag of fertilizer from the nearest input dealer:

* `fmkt_pri_bas`: basal fertlizer, (compound D). One of the most used fertlizer (cf:http://www.fao.org/3/a0395e/a0395e0a.htm).
* `fmkt_pri_top`: top-dressing fertilizer (urea).
* `fmkt_mkt_trp`: transport.

#### Other variables

Those variables do not relate to a fix amount of fertilizer.
They all start with the prefix `price`. The middle part is then a code for the type of input:

| code   |               fertilizer              |
|--------|:-------------------------------------:|
| fert1  |                   SA                  |
| fert2  |                  DAP                  |
| fert3  |                  CAN                  |
| fert4  |                  Urea                 |
| fert5  |                  TSP                  |
| fert6  |             MinjinguMazao             |
| fert7  |             YaraMilaCereal            |
| fert8  |               YaraAmidas              |
| fert9  |             YaraMilaWinner            |
| fert10 |                 other                 |
| oinp1  | pesticide (for field use not storage) |
| oinp2  |               herbicide               |
| oinp4  |               fungicide               |

Lastly, a suffix indicates the type of variable:

| suffix |   variable   |
|--------|:------------:|
| tsh    | price in Tsh |
| amt    |    amount    |
| unit   |     unit     |

Example: `price_fert2_amt` amount of DAP, given in the unit indicated in `price_fert2_unit` and corresponding to the price in Tsh in `price_fert2_tsh`.

### Subset main community data.frame

```{r}
cmty_fert <- cmty %>% 
  select(cmty_index, 
         starts_with("fmkt"),
         matches("^price_oinp[^3]"),
         starts_with("price_fert"))
```

Any variables with negative values ?

```{r}
cmty_fert %>%
  select(-cmty_index) %>%
  any_negative()
```

The value -98 seems to be the recurring place holder for missing value, sometimes it is also mislabelled as 98...

```{r}
cmty_fert[c(42,54), c(1, 5, 14)]
```

Replace those place holder values by NA.
```{r}
id_vars <- grep("cmty_index", colnames(cmty_fert), invert = TRUE)
err_to_NA <- which(cmty_fert[, id_vars ] <= 0 | cmty_fert[, id_vars] == 98, arr.ind = TRUE)
cmty_fert[, id_vars][err_to_NA] <- NA
```

### Calculate price in Tsh/kg for each fertilizer

#### Transport price per kg

Transport price is assume to scale linearly with the amount being transported and is calculated by simply dividing the cost of transport of a 50kg fetlizer bag by 50.

```{r}
cmty_fert <- cmty_fert %>% mutate(transp_price = fmkt_mkt_trp / 50)
```

```{r, warning = FALSE}
plot_var(cmty_fert, var = transp_price, plot.unit = "Tsh/kg")
```

```{r}
cmty_fert$transp_price[cmty_fert$transp_price > 200] <- NA
```

```{r, warning = FALSE}
plot_var(cmty_fert, var = transp_price, plot.unit = "Tsh/kg")
```

#### Price Tsh/kg

Per unit category, per fertilizer type, calculate price in Tsh/kg as:

$Price_{(Tsh.kg^{-1})} = \frac{Price_{(Tsh)}}{amount_{(kg)}} + transport_{(Tsh.kg^{-1})}$

```{r}
fertilizer <- c("SA", "DAP", "CAN", "Urea", "TSP",
                "MinjinguMazao", "YaraMilaCereal", "YaraAmidas", "YaraMilaWinner", "other")
pesticide <- c("pesticide", "herbicide", NA, "fungicide")
units <- c(1, 2, 3, 5, 6, 99)
```

```{r}
# Fertilizer
cmty_fert <- price_input(df = cmty_fert,
                         input_type = "fert",
                         input_id = 1:10,
                         input_name = fertilizer,
                         units = units)
# Pesticide 
cmty_fert <- price_input(df = cmty_fert,
                         input_type = "oinp",
                         input_id = c(1, 2, 4),
                         input_name = pesticide,
                         units = units)
```


#### Keep only transformed variables 

```{r}
cmty_fert <- cmty_fert %>%
  select(cmty_index, fmkt_pri_bas, fmkt_pri_top, transp_price,
         paste0("price_", pesticide[!is.na(pesticide)]),
         paste0("price_", fertilizer))
```

Rename standard bag variables (`fmkt`)

```{r}
colnames(cmty_fert)[2:3] <- c("price_basfert", "price_topfert")
```

Convert those in price per kg

```{r}
cmty_fert <- cmty_fert %>% 
  mutate(price_basfert = price_basfert / 50,
         price_topfert = price_topfert / 50)
```


#### Visual exploration

```{r, warning = FALSE}
plot_var_facet(cmty_fert, 
               vars = c("price_basfert", "price_topfert"), 
               plot.unit = "Tsh/kg", 
               title = "Price base and top fertilizers")
```

```{r, warning = FALSE}
var_fert <- map(fertilizer, ~ grep(.x, colnames(cmty_fert), value = TRUE)) %>% unlist()
plot_var_facet(cmty_fert, 
               vars = var_fert, 
               plot.unit = "Tsh/kg", 
               title = "Price fertilizers")
```

`price_TSP`, all NA ?: ``r all(is.na(cmty_fert$price_TSP))``.

```{r}
cmty %>%
  select(starts_with("price_fert5")) %>% 
  head()
```

```{r}
table(cmty$price_fert5_tsh)
```

```{r, warning = FALSE}
var_pest <- map_chr(pesticide[!is.na(pesticide)], 
                    ~ grep(.x, colnames(cmty_fert), value = TRUE))
plot_var_facet(cmty_fert, 
               vars = var_pest, 
               plot.unit = "Tsh/kg", 
               title = "Price inputs other than fertilizers")
```


### Price of N, P and K in $Tsh.kg^{-1}$

We now calculate the price of fertilizer by isotopes, namely N, P and K. This price corresponds to the price of a given raw isotope in $Tsh.kg^{-1}$. We do not take into account the difference in terms of availability of N, P and K per fertilizer type. 


Modified input names in dedicatd dictionary (`dics$inp`) so that they match community data names.

```{r}
str_remove(dics$inp$input, "(\\(.*\\))") %>%
  str_remove(" ") -> dics$inp$input_mod
```

Temporarily transform `cmty_fert` to long format to be able to join it with `dics$inp` and perform calculations.

```{r}
cmty_fert_l <- cmty_fert %>% 
  select(cmty_index, starts_with("price")) %>%
  pivot_longer(cols = str_subset(colnames(cmty_fert), "^price"), 
               names_to = "input", 
               values_to = "price")
cmty_fert_l$input <- str_extract(cmty_fert_l$input, "(?<=price_).*")
```

Perform calculation and transform back to wider format `cmty_fert`.

```{r}
cmty_fert_l <- cmty_fert_l %>% full_join(dics$inp, 
                                         by = c("input" = "input_mod"), 
                                         suffix = c(".x", ".dic"))
```

Define function to divide fertilizer price by concentration only if the concentration of a given isotope is different from 0.

```{r}
isotope_price <- function(price, conc){
  map2_dbl(.x = price, .y = conc, function(.x, .y){ ifelse(.y != 0, .x/.y, NA)})
}
```

Calculate price per isotope.

```{r}
cmty_fert_l <- cmty_fert_l %>% mutate(N_price = isotope_price(price, Nconc), 
                                      P_price = isotope_price(price, Pconc),
                                      K_price = isotope_price(price, Kconc))
```

Convert to wider format.

```{r}
cmty_fert_NPK <- cmty_fert_l %>% 
  pivot_wider(names_from = "input", values_from = 'price')%>%
  select(cmty_index, input.dic, matches("^[A-Z]_price$")) %>%
  rename(input_type = input.dic)
```

Delete rows without any input or isotope price.

```{r}
cmty_fert_NPK <- cmty_fert_NPK[rowSums(is.na(cmty_fert_NPK)) < 3, ]
```

Plot results.

```{r, warning = FALSE}
dNPK <- pivot_longer(cmty_fert_NPK, cols = paste0(c("N", "P", "K"), "_price"),
                     names_to = "category",
                     values_to = "value")
dNPK$category <- factor(dNPK$category, levels = c("N_price", "P_price", "K_price"))
plot_var(dNPK, var = value, plot.unit = plot.unit)+
  facet_wrap(. ~ category)+
  xlab("")+
  ggtitle("Fertilizer price per isotope: N, P and K")
```


## Labour cost

Cost of hiring one adult man for one labour day.

**Labour variables:**

* prefix: `labprc`.
* code: `clearing`, `hoeing`, `planting`, `tillplnt`, `fertapp`, `weeding`, `pestapp`, `harvesting`, `shelling`, pretty transparent except `tillplant` = tilling and planting together.
* suffix: `price`.

### Subset main community data.frame

```{r}
cmty_labour <- cmty %>% 
  select(cmty_index, matches("^labprc_.*_price$"))
```

Get rid of variable prefix.

```{r}
names(cmty_labour) <- str_replace(names(cmty_labour), "labprc_", "")
```

Get all labour variable names.

```{r}
lab_vars <- grep("_price$", colnames(cmty_labour), value = TRUE)
```

Any variables with negative values ?

```{r}
cmty_labour[, lab_vars] %>% any_negative()
```

Replace negative values by NA.

```{r}
cmty_labour[, lab_vars] <- modify(cmty_labour[, lab_vars], ~ ifelse(.x < 0, NA, .x))
```

### Visual exploration

```{r, warning = FALSE, message = FALSE}
plot_var_facet(cmty_labour, 
               vars = lab_vars, 
               plot.unit = "Tsh / work day", 
               title = "Community labour prices") +
  scale_y_continuous(labels = scales::label_number_si())
```


```{r}
cmty_labour <- modify_at(.x = cmty_labour,
                         .at = lab_vars,
                         .f = outlier_to_NA,
                         quantile = 0.95)
```

```{r, warning = FALSE, message = FALSE}
plot_var_facet(cmty_labour, 
               vars = lab_vars, 
               plot.unit = "Tsh / work day", 
               title = "Community labour prices") +
  scale_y_continuous(labels = scales::label_number_si())
```


## Merge 

Merge together.

```{r}
community <- reduce(list(cmty_info, cmty_seed, cmty_fert, cmty_labour),
                    inner_join,
                    by = "cmty_index")
```

Save intermediate dataset.

```{r}
write_csv(x = community,
          path = here("./data/extracted/community.csv"))
```

Save separate dataset with fertilizers isotope prices.

```{r}
community_NPK <- inner_join(cmty_info, cmty_fert_NPK, by = "cmty_index")
write_csv(x = community_NPK,
          path = here("./data/extracted/community_NPK.csv"))
```



## Recap 

### Community general dataset

```{r}
read.csv(here("./docs/tables/community_recap_tab.csv")) %>%
  knitr::kable()
```

### Fertilizer isotope prices

```{r}
read.csv(here("./docs/tables/community_NPK_recap_tab.csv")) %>%
  knitr::kable()
```
