---
title: "TAMASA - data cleaning summary"
author: "Antoine Languillaume"
date: "12/16/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
source(here("./scripts/plot_funcs.R"))
```


```{r, message = FALSE}
d <- read_csv(here("./data/extracted/household.csv"), na = "NA")
```

## Paul meta analysis experiments

```{r, echo = FALSE}
magick::image_read(here("./docs/schemes/experiment_sites.png")) %>%
  magick::image_resize("200x200%")
```

## Where I started

- Taking other Hilde.
- Main problem: could not understand her work.
- Some things done manually and not reproducible.

## Methods brief

- R markdown
- Git via gitlab


A necessity and not a luxury


## Tanzania cropping season

```{r, echo = FALSE}
magick::image_read(here("./docs/schemes/growing_seasons.png")) %>%
  magick::image_resize("60x60%")
```

## Enumeration data

```{r, echo = FALSE}
enum_date <- d %>% 
  select(enum_date, zone) %>%
  group_by(enum_date, zone) %>%
  summarise(n = table(enum_date)) %>%
  rename(date = enum_date)
enum_date$n <- as.numeric(enum_date$n)
enum_date$date <- lubridate::as_date(enum_date$date)
enum_date <- enum_date[lubridate::year(enum_date$date) == 2017, ]
```

```{r, echo = FALSE}
ggplot(enum_date)+
  aes(x = date, y = n, group = 1, color = zone) +
  geom_point()+ 
  geom_segment(aes(xend = date, yend = 0))+
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d")+
  ylab("Number of household interviewed")+
  ggtitle("Enumeration date TAMASA survey")
```


## Inherant noise in the data


Variables not measured but estimated by farmer:

- plot size
- yield (most of the time crop not harvested) 

## Labour data

```{r, warning = FALSE}
labour_harvest_plot(d[!is.na(d$cropsys), ])
```


## Labour data vs Paul 

```{r, warning = FALSE}
labour_harvest_plot(d[!is.na(d$cropsys), ])+
     geom_abline(slope = 0, intercept = 3, color = "red")
```


## Core of the problem

Interpretation of the question:

**actual time spent versus time range**


## Esther experience

In Eshter's plots it takes 5 men to harvest 1 acre / day of maize, with 8 hours worked per day.

```{r, echo = TRUE}
acres_day = 1/5
ha_day = acres_day / 2.47
day_ha = 1 / ha_day
ton_ha = seq(1.5, 3.5, by = 0.1)
```

```{r}
l = list(data.frame(dt_8 = day_ha / ton_ha),
         data.frame(dtr_8 = day_ha / d$yield[d$yield > 1]),
         data.frame(dtaps_8 = d$tlab_h_dt[d$crop == "maize"]))

dt = plyr::rbind.fill(l)

dt = dt %>% mutate(dt_h = dt_8 * 2, 
                   dtr_h = dtr_8 * 2,
                   dtaps_h = dtaps_8)
```

```{r}
dt = pivot_longer(dt, cols = grep("dt", colnames(dt)),
                  names_to = "type",
                  values_to = "value")
dt = separate(dt, 
              col = type, 
              into = c("source", "hours_per_day"),
              sep = "_")
```

```{r}
dt$source <- dt$source %>% 
  fct_recode("Esther" = "dt",
             "Esther - TAMASA yield" = "dtr",
             "TAMASA - raw" = "dtaps") %>%
  fct_relevel("Esther", "Esther - TAMASA yield", "TAMASA - raw")

dt$hours_per_day <- dt$hours_per_day %>% fct_recode( "8h / day" = "8", "4h /day" = "h")
```


## Plot Esther experience

```{r}
dt %>%
  filter(source != "TAMASA - raw") %>%
ggplot()+
  aes(y = value, x = source)+
  geom_jitter(alpha = 0.5, na.rm = TRUE)+
  geom_boxplot(na.rm = TRUE, alpha = 0.5)+
  geom_abline(slope = 0, intercept = 3, color = "red")+
  ylab(expression(days.ha^-1))+
  facet_wrap(. ~ hours_per_day)
```


## Plot Esther experience vs TAMASA

```{r, fig.width = 8}
dt %>%
ggplot()+
  aes(y = value, x = source)+
  geom_jitter(alpha = 0.5, na.rm = TRUE)+
  geom_boxplot(na.rm = TRUE, alpha = 0.5)+
  geom_abline(slope = 0, intercept = 3, color = "red")+
  ylab(expression(days.ha^-1))+
  facet_wrap(. ~ hours_per_day)
```



## Matrix plot per cropping system

```{r}
pal <- RColorBrewer::brewer.pal(n = 10, "RdYlGn")
pal <- pal[length(pal):1]
d %>% 
  filter(zone == "N") %>%
  select(hh_index, zone, tlab_h_ha, hlab_h_ha, flab_h_ha, plot_area, yield, cropsys, crop) %>%
  pivot_longer(cols = str_subset(colnames(.), "lab"),
               names_to = "bar", 
               values_to = "value") %>% 
  separate(bar, into = c("lab_cat", "task", "area"), sep = "\\_") %>%
  ggplot()+
  aes(y = value, x = plot_area, color = yield)+
  geom_point(na.rm = TRUE)+
  scale_color_gradientn(colours = pal )+
  ylab(expression(days.ha^-1))+
  facet_grid(lab_cat ~ cropsys+crop)+
  theme_bw()+
  theme(panel.grid = element_blank())
```


## Matrix Plot

```{r, message = FALSE, warning = FALSE, fig.dim = c(7, 5), fig.align = 'center'}
d %>% select(tlab_h_ha, tlab_h_dt, plot_area, yield) %>% GGally::scatmat()
```

```{r, }
# d %>% 
#   filter(zone == "N") %>%
#   select(hh_index, plot_id, tlab_h_ha, tlab_h_dt, yield, crop, cropsys) %>%
#   group_by(cropsys, crop) %>%
#   summarise(mean_harvest_lab_ha = mean(tlab_h_ha, na.rm = TRUE),
#             mean_harvest_lab_dt = mean(tlab_h_dt, na.rm = TRUE))
```


## What went wrong

- Hidden layers of complexity
- Should have contacted Jordan from the start


# What's next