---
title: "TAMASA APS17 - Jordan Focal Plot data"
author: "Antoine Languillaume"
date: "09/03/20"
output_type: html_document
---

```{r, message = FALSE}
library(tidyverse)
library(haven)
library(here)
library(broom)
library(plotly)
```

### Read in data

```{r}
fp <- haven::read_stata(here("./data/Jordan/TZAPS17_fp.dta"))
area <- fp %>% select(matches("area"))
```

`area` variables meta data:

* `area_est`: plot area estimated by farmer in __acres__.
* `area_gps`: gps reading in $m^{2}$.
* `area_gps_calc`: plot area in acre calculated from GPS measurement.
* `area_gps_conf`:	Was the plot polygon saved by the enumerator in his device.

### Check reliability of the conversion.

Were $m^{2}$ properly converted to acres ?

```{r}
area$area_calc_check <- area$area_gps / 4046.86
range(na.omit(area$area_gps_calc - area$area_calc_check))
```
The answer is yes as the difference between the original variable `area_gps_calc` and the variable we computed `area_calc_check` amounts only to computer rounding errors.

### Plotting data

Let's assess the precision of the farmers' estimate by fitting a simple linear regression.

```{r, warning = FALSE}
area <- area %>% filter(!is.na(area_gps_calc))
fit <- lm(area_gps_calc ~ area_est, data = area)
pred <- as.data.frame(predict(fit, interval = "prediction", level = 0.95))
adjr2 <- glance(fit) %>% select(matches("adj.r.squared")) %>%  deframe() %>% round(2)
```

**The adjusted $R^2$ for our model is `r adjr2`.** In comparison Marloes found $R^2$'s of 0.55 and 0.63 for the two sites she studied in Ethiopia.


#### Interactive plot

```{r, warning = FALSE, fig.width = 10}
p <- plot_ly(x = ~ area_est, 
        y = ~ area_gps_calc, 
        type = "scatter",
        mode = "markers",
        data = area,
        name = "raw data")
p %>% 
  add_lines(x = ~ area_est, y = fitted(fit), name = "lm") %>%
  add_ribbons(x = ~ area_est, ymin = pred$lwr, ymax = pred$upr, name = "lm pred 95% CI")
```




